### 2009-12-07 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

\chapter{Conclusions and Future Works}

@@

@@

Taking into account the software architecture provided by NetBEAMS, as well as
the requirements for data collection from the RTC research group, this work
shows a novice way to persist data collected from sensor devices by using a
schema-less data model, whose query capabilities are based upon programming
languages, in contrast to the use of the Structured Query Language, SQL. The
experiment results showed that the implementation of a persistence layer with
mongoDB can be used to persist data at its maximum rate of one sample per
minute. However, performance can be improved by using techniques such as
database sharding, since it offers a direct implementation capability to
Data-Centric Storage approach. This approach decreases the amount of time to
search data since it distributes the data instances across different cluster
nodes.

There are many different directions for future works. First, the Data-Centric
implementation performance may be extremely improved by using the technique
called MapReduce \cite{map-reduce}. This approach allows mongoDB to search
different machines in parallel whenever the data requested is located in
different database shard hosts. In addition to that, the use techniques to
cluster collected data before they are inserted into the database may be
important as shown in \cite{sn-time-series}. One of the key issues in the
implemented solution is that any measurement of data is considered to be
persisted without further analysis.

All in all, the DSP Data Persistence component implemented for the NetBEAMS
solution to SF-BEAMS can be used in any network server. The solution for data
persistence is a novice way to save data into the database, provided by
mongoDB's ability to cope with the uncertainty of any properties collected from
any sensor specification. Furthermore, the solution also provides a good
support to the provenance-related data.

>>>>

\chapter{Conclusions}

@@

As seen in the state of the art, different approaches can be done in order to
evaluate data persistence for nsensor networks. T

@@

Taking into account the software architecture provided by NetBEAMS, as well as
the requirements for data collection from the RTC research group, this work
shows a novice way to persist data collected from sensor devices by using a
schema-less data model, whose query capabilities are based upon programming
languages, in contrast to the use of the Structured Query Language, SQL. The
experiment results showed that the implementation of a persistence layer with
mongoDB can be used to persist data at its maximum rate of one sample per
minute. However, the evaluation of mongoDB was not used to implement a
Data-Centric Storage approach, which can decreases the amount of time to search
data since the process uses parallel search.

All in all, the DSP Data Persistence component implemented for the NetBEAMS
solution to SF-BEAMS can be used in any network server. The solution for data
persistence is a novice way to save data into the database, provided by
mongoDB's ability to cope with the uncertainty of any properties collected from
any sensor specification. Furthermore, the solution also provides a good
support to the provenance-related data as the schema-less data model gives the
ability to .

\chapter{Future Works}

0. Todo list for this work
1. Database System Problems
2. Software Engineering Problems
3. Applications
 - plot graphs of collected,

There are many different directions for future works. First, the
implementation of a Data-Centric Storage approach can help the system scale
horizontaly by using the mongoDB's database shards. In this way, this approach
can reduce the retrieval time by focusing on centralized selection of data over
a given database shard. Moreover, the overall search through the collected data
attributes can be drastically decreased by parallel search over the database
shards using the MapReduce \cite{map-reduce} tools. At this time of
development, the MapReduce API offered by mongoDB is on its alpha version, and
therefore, presenting lots of bugs. Figure
\ref{fig:future-works-data-centric-map-reduce} shows an architectural idea of
the deployment of a Data-Centric storage approach. When searching for specific
document attributes, the process is sent in parallel to each of the database
shards, and the result is collected by the cluster head. After calculating the
final result, mongoDB server returns the relevant result to the user.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/future-works-data-centric-map-reduce}
  \caption{Data-Centric approach using MapReduce to search accross different
  database shards}
  \label{fig:future-works-data-centric-map-reduce}
\end{figure}

In addition to infrastructural, the use techniques for scheduling data
gathering from the SF-BEAMS sensor networks can be used to decrease the data
load into the system. While the period of times used by the RTC seems
to cover the basic requirements of data, the data storage may be
seen as overloaded with repeated data. Since environmental conditions
does not change drastically over time, the use of data scheduling and
data clustering approaches can be used. First, considering that SF-BEAMS
contains a cluster of sensors, the data load could be decreased by using
techniques such as clustering the data in the data sink before they are
persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
the knowledge about the data can only be achieved in the data sink. In this
way, one suggestion is the implementation of a DSP Data Clustering component
that is responsible for the data clustering before they are inserted into the
database. Similarly, the use of schedulers on the sensors could also be used to
decrease the amount of data to be sent to the data sink, as seen at the survey
\cite{sn-scheduling-survey}.

Finally, taking into account the use of the database system to complement
NetBEAMS execution, the design of event-based applications could add different
functionalities to the management capabilities of the sensor networks. For
instance, the YSI sonde collected data carries the information regarding the
battery life of the device, seen at the collected attribute 
``observation.Battery''. A new DSP Component responsible for observing these
attributes could define the threshold of the event of low-battery. With an
event-based application in mind, events regarding the SF-BEAMS network could be
better managed through NetBEAMS where scheduled sites visits would depend on
such events to happen and, thus, avoiding unecessary operational costs.

### 2009-12-07 docs/marcello-dissertation/thesis/preamble.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

\singlespace{
\myname \\
San Francisco, California \\  
\thisyear \\
}

\end{center}

@@

%\doublespacing{\noindent
\onehalfspacing{\noindent      
Sensor Networks are becoming more and more important to different science
communities bacause of what it produces: the raw data for different studies.
In order to make use the collected data, researchers may have to dissect the
characteristics of the sensor network in question, regarding different
properties such as the purpose and location of the observed data, as well as
how the data is described. In this way, this work first proposes a set of
taxonomies that can be used to study and classify a given instance of sensor
network and its produced data. Then, based on the classes of these taxonomies, 
this work selected a Document-Oriented Database system as a result of an
empirical evaluation of the requirements for a new persistence layer for 
NetBEAMS, the case study used in this work. For this reason, the design and
implementation of a new data persistence component for NetBEAMS show how
schema-less data models help decreasing the complexity related to the database
layer, not yet reported by the state of the art. Furthermore, this work also
proposes the use of Database Shards in order support the implementation of
Data-Centric sensor networks. 
}
\vspace*{\fill}

>>>>

\singlespace{
\myname \\
San Francisco, California \\  
\thisyear \\
}
\end{center}

@@

%\doublespacing{\noindent
\onehalfspacing{\noindent      
Sensor Networks are becoming more and more important to different science
communities bacause of what they produce: the raw data for different studies.
In order to make use the collected data, researchers may have to dissect the
characteristics of the sensor network in question, regarding different
properties such as the purpose and location of the observed data, as well as
how the data is described. For this reason, this work proposes a set of data
persistence taxonomies based on the state of the art in data persistence for
sensor networks, which can be used to classify the infrastructure of an
existing sensor network and the nature of its produced raw data. In order to
evaluate the proposed taxonomies, this work designs and implements a data
persistence component for NetBEAMS, a modular sensor network infrastructure
developed to improve the operation of the SF-BEAMS environmental Sensor
Network. In this way, based on empirical analysis of the proposed taxonomies
and the requirements of a persistence layer for the case study, this work
selected mongoDB, a schema-less document-oriented database system intead of
the traditional use of relational databases. As a result of the experiments
conducted with mongoDB, this work concluded that schema-less data models can
help decreasing the complexity related to the data persistence layer for
sensor networks with external or data-centric storage mechanism, and whose
purpose is data archival. Finally, this work proposes different future works
such a data-centric approach using Database Shards to store the collected
data and Map Reduce to execute parallel queries over the shards.}
\vspace*{\fill}
